Are you normal? A new projection pursuit index to assess a sample against a multivariate null distribution

### 2024-07-03 paper.qmd
language smoothing

Another application would compare outlying points to a reference normal distribution that is derived using robust methods on the original data. To illustrate this we work on the weather data example from @filzmoser2018. The data contains 16 numerical measurements that are averages across the three summer months June, July and August, these are provided for 68 years (1955 - 2022), see @filzmoser2018 for more details.

@@

We first apply the new anomaly index to the full dataset, such that the final projection will identify a compromise and provide a global picture of where the outlying points differ from the normal distribution.

@@

>>>>

This example is motivated by the weather data example from @filzmoser2018. We illustrate using the anomaly index to  compare potential outliers with a reference normal distribution that is derived using robust methods on the original data. The data contains 16 numerical measurements that are averages across the three summer months June, July and August, these are provided for 68 years (1955 - 2022), see @filzmoser2018 for more details.

@@

The new anomaly index is first applied to the full dataset, such that the final projection will identify a compromise and provide a global picture of where the outlying points differ from the normal distribution.

@@

# adding labels and running grand tour to better track
# why there might be a missmatch
animate_xy(X,
           ellipse=as.matrix(Sigma), ellc = size_weather,
           axes="off", col = as.factor(cluster_labels),
           palette = "Plasma", half_range = 5,
           center = FALSE, rescale = FALSE, obs_labels = rownames(X))
